---
title: "b) A brief introduction to matrix models"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{b) A brief introduction to matrix models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(popcycles)
```


An excellent introduction to matrix population models can be found here:

Stevens, H.  2020.  A primer of ecology in R
Chapter 4 Density-independent Demography
Sections 4.1 to 4.6
https://hankstevens.github.io/Primer-of-Ecology/DID.html

Key concepts:

* life cycle graph
* matrices and matrix math
* life stages
* matrix representation of life cycles
* population projection
* population growth rate
* population growth rate from eigenanalysis
* sensitivity and elasticity analysis
* what's missing from this model: spatial structure, density dependence


## Example: Gopher Tortise

Desert tortoise "medium-high" fertility matrix from Doak et al. 1994  Used for examples in Caswell 2002 book, Wisdom et al, and elsewhere.  Results reported in Table 6 of Doak et al. 1994.

###  Data

Data for testing / demonstrating functions.  Could add this to help files for the functionss.

```{r}
         
tortoise <- matrix(data = c(0.000,0.000,0.000,0.000,0.000,1.300,1.980,2.57,
0.716,0.567,0.000,0.000,0.000,0.000,0.000,0.00,
0.000,0.149,0.567,0.000,0.000,0.000,0.000,0.00,
0.000,0.000,0.149,0.604,0.000,0.000,0.000,0.00,
0.000,0.000,0.000,0.235,0.560,0.000,0.000,0.00,
0.000,0.000,0.000,0.000,0.225,0.678,0.000,0.00,
0.000,0.000,0.000,0.000,0.000,0.249,0.851,0.00,
0.000,0.000,0.000,0.000,0.000,0.000,0.016,0.86),
byrow = T,nrow = 8)

n <- c("yearling",  "juv1",  "juv2",  "imm1",  "imm2", "subadult", "adult1", "adult2")
rownames(tortoise) <- n
colnames(tortoise) <- n
```


Test calc_lam
```{r}

calc_lam(tortoise) #Matches "Average" - "All" result, Table 1, Doak et al 1994.
```


## Sensitivities

```{r}

max(calc_S(tortoise)$E) # matches sensitivity reported in Caswell 2002 pg 256


```


Note: elasticities sum to 1.
```{r}
sum(calc_S(tortoise)$E)
```

