---
title: "m) Notes on other papers from ReScience-C"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{m) Notes on other papers from ReScience-C}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(popcycles)
```


## Example paper: [Re] Chaos in a Three-Species Food Chain


### intro - real lit review

One of the most important components of an ecological community is its food web.
Early models of population dynamics generally only considered interactions between
two species (e.g., Canale1; Rosenzweig and MacArthur2). However, in nature, food webs
wherein only two species influence the behaviour of the network are quite uncommon
– most networks are far more complex [3]. Therefore, several researchers asserted that
every food web study should involve at least three species in order to capture that complexity
[4, 5].
At first, the core interest of food web research was in equilibrium analysis, partly based
on the assumption that communities observed in nature represent a stable equilibrium
state [6]. However, complex dynamics arise in a system with three or more species, and
these are better described as chaos rather than stable equilibrium. The simplest definition
of chaos is the extreme sensitivity of a system to its initial conditions [6]. Hastings
and Powell3, who studied chaos in a continuous‐time model of a food web including
three species, contributed considerably to the significance and understanding of this
subject. This study led to many others on food webs dynamics and chaos, which reinforced
the importance of chaos in ecological modelling [7, 8].
In this article, we provide a replication of the model from Hastings and Powell3 using the
same equations and parameter values. Replicating a classic study and model like this
one is important for many reasons: 1) we can compare the original results with ones
obtained using current technologies, ensuring that they can still be reproduced; 2) we
can provide a publicly available and reproducible version of the codes that recreate the
model. We were able to successfully replicate all figures in the original article using Julia
v1.3.1. All code used to replicate the original model is available alongside this article
(https://github.com/BIO6032/2019_replication_HastingsPowell_1991).

## Example paper: [Re] Insect natural enemies as regulating factors

has real lit review

## Example paper: [Re] On the coexistence of specialists and generalists

terse

## Example paper: [Re] Least-cost modelling on irregular landscape graphs

terse
